One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-10 paper/results_3d.tex
intro draft completed
begun summary, updated 3d figs

\section{Three-dimensional simulations}
In this section we briefly review 3D simulations carried 
out using ZEUS-MP and PLUTO. The main purpose is to check 
the above results with different numerical codes, and validate  
the 2D approximation.

@@

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs. We also show results from strictly
isothermal discs ($q=0$), which clearly display no growth. %why no high m
                                %modes as in 2D for iso disc?
                                %resolution?
Results from ZEUS-MP appears to be off-set from that obtained using
PLUTO. We suspect this may be due to larger numerical noise in the
ZEUS-MP code. Thus the $m=1$ spiral grows to a larger amplitude by the
end of the run. However, we measure similar growth rates,
\begin{align*}
&\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
&\gamma \simeq 0.0075\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations above. This may be due to the
low resolutions adopted. 
%growth rate does not depend on pert amp

@@

\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.33]{figures/polarxy_dens015_z0}
      \includegraphics[scale=0.33,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/polarxy_dens015_zmax}  
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.33]{figures/pdisk_021_z0}
      \includegraphics[scale=0.33,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/pdisk_021_zmax}
    }
  \end{center}
  \caption{Three-dimensional simulations using the ZEUS-MP (top) and 
    PLUTO (bottom) codes. The non-axisymmetric density at the midplane (left) and
    two scale-heights above the midplane (right) is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height  
    from the midplane.\label{3d_prelim}}   
\end{figure}

>>>>

\section{Three-dimensional simulations}\label{results3d}
In this section we briefly review 3D simulations carried 
out using ZEUS-MP and PLUTO. The main purpose is to check 
the above results with different numerical codes, and validate  
the 2D approximation.

@@

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs. We also show results from strictly
isothermal discs ($q=0$), which clearly display no growth. %why no high m
                                %modes as in 2D for iso disc?
                                %resolution?
Results from ZEUS-MP appears to be off-set from that obtained using
PLUTO. We suspect this may be due to larger numerical noise in the
ZEUS-MP code. Thus the $m=1$ spiral grows to a larger amplitude by the
end of the run. However, we measure similar growth rates,
\begin{align*}
&\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
&\gamma \simeq 0.0084\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations above. This may be due to the
low resolutions adopted. 
%growth rate does not depend on pert amp

@@

\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.33]{figures/polarxy_dens015_z0}
      \includegraphics[scale=0.33,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/polarxy_dens015_zmax}  
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.33]{figures/pdisk_020_z0}
      \includegraphics[scale=0.33,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/pdisk_020_zmax}
    }
  \end{center}
  \caption{Three-dimensional simulations using the ZEUS-MP (top) and 
    PLUTO (bottom) codes. The non-axisymmetric density at the midplane (left) and
    two scale-heights above the midplane (right) is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height  
    from the midplane.\label{3d_prelim}}   
\end{figure}

### 2014-11-10 paper/summary.tex
begun summary, updated 3d figs

\section{Summary and conclusions}\label{summary}

@@

>>>>

\section{Summary and conclusions}\label{summary}
In this paper, we carried out direct hydrodynamic simulations of 
radially structured, self-gravitating locally isothermal discs. 
We find such systems can be unstable to perturbations with azimuthal
wavenumber $m=1$, which lead to the development of one-arm 
trailing spirals that persist for at least $O(10^2)$ orbits. The
spiral pattern speed is smaller than the local disc rotation, making
it a low-frequency mode that is almost stationary.

According to previous studies \citep{adams89}, our disc models
should not support $m=1$ linear spiral instabilities. This is because
our discs are not very massive ($M_d\lesssim 0.1M_*$) and, more
importantly, we have neglected the motion of the central star. The 
latter was shown to be important for one-arm spiral modes to grow  
\citep{shu90}.

In order to explain our numerical results, we appealed to  
angular momentum conservation within linear theory. We find that a  
forced temperature gradient introduces a torque on linear
perturbations. This torque

%works in 2D/3D
%forced gradient feedback 
%neutral mode destabilized

@@

%numerical simulations to include cooling/heating
%non-sg sims, Be stars, eccentric discs
%linear theory: explicit calculations
